\Title

### 2016-08-17 manuscript/manuscript.tex
Replace cross-validation with validation using seismological data
Fixes #29

The Mohorovičić discontinuity (or Moho) that marks the transition from the
crust to the mantle, is studied almost exclusively through indirect geophysical
methods.
The two main geophysical methods used to estimate the depth of the Moho are
seismology, with both natural and controlled sources, and gravimetry.
With the advent of satellite gravimetry missions like GRACE and GOCE,
gravity-derived crustal models can be produced in regional or global scales
\citep[e.g. ][]{reguzzoni2013,vandermeijde2013,vandermeijde2015}.
New spherical harmonic gravity models that use these satellite observation,
like GOCO5S \citep{mayer-guerr2015}, provide almost homogeneous data coverage
in difficult to access regions traditionally poor in terrestrial data.
An example is South America, where seismologic and terrestrial gravity data
are traditionally concentrated around urban centers and coastal areas,
resulting in large areas (e.g., forests and mountains) devoid of data.

@@

We present a non-linear gravity inversion to estimate the Moho depth in
a spherical Earth approximation.
Our method is based on the \citet{silva2014} Gauss-Newton formulation of the
method of \citet{bott1960}.
We use tesseroids to discretize the anomalous Moho and the adaptive
discretization algorithm of \citet{uieda2016} for the forward modeling.
The stability of the inversion is achieved through smoothness regularization.
In order to maintain the computational efficiency of Bott's method,
we exploit the sparse nature of all matrices involved in the computations.
We employ a variant of GCV known as hold-out cross-validation \citep{kim2009}
to estimate the regularization parameter.
Additionally, we estimate the density-contrast and reference level
simultaneously in a second cross-validation.
Similarly to \citet{silva2006} and \citet{martins2010}, this cross-validation
procedure uses knowledge of the Moho depths at certain points.
Finally, we apply the proposed method to estimate the Moho depth for South
America using gravity data from the GOCO5S model \citep{mayer-guerr2015} and
the seismological data of \citet{assumpcao2013a}.

@@

We estimate these hyper-parameters in two steps.
First, we assume fixed values for
$z_{ref}$ and $\Delta\rho$ and perform a cross-validation procedure
\citep{hansen1992} to estimate an optimal value for $\mu$.
Our investigations suggest that the optimal value of $\mu$ does not depend on
the particular values of $z_{ref}$ and $\Delta\rho$ used.
Second, we use the estimated $\mu$ to perform a second cross-validation
procedure to estimate $z_{ref}$ and $\Delta\rho$.
The outcome of both cross-validations is not only the values of the three
hyper-parameters but the final estimated Moho depths.

@@

The depth of the Normal Earth Moho ($z_{ref}$)
and the density-contrast of the anomalous Moho ($\Delta\rho$)
are other hyper-parameters of the inversion.
That is, their value influences the final solution
but they are not estimated during the inversion.
Both hyper-parameters cannot be determined from the gravity data alone.
Estimating $z_{ref}$ and $\Delta\rho$ requires
information that is independent of the gravity data,
such as knowledge of the parameters at certain points.
This information can be used in a manner similar to
the cross-validation described in the previous section.
In this study, we use point estimates of the Moho depth
to determine the optimal values of $z_{ref}$ and $\Delta\rho$.
These points will generally come from seismologic studies,
like receiver functions, surface wave dispersion, and deep refraction
experiments.

@@

We test and illustrate the proposed inversion method
by applying it to two noise-corrupted synthetic data sets.
The first one is generated by a simple Moho model simulating the transition
from a thicker continental crust to a thinner oceanic crust.
This application uses cross-validation to estimate the regularizing parameter
($\mu$) while assuming that the anomalous Moho density-contrast ($\Delta\rho$)
and the Normal Earth Moho depth ($z_{ref}$) are known quantities.
This first test is simplified in order to investigate solely
the efficiency of the inversion and
the cross-validation procedure to estimate $\mu$.
The second data set is generated by a more complex model derived from
the South American portion of the global CRUST1.0 model \citep{laske2013}.
This application uses cross-validation to estimate all three hyper-parameters:
$\mu$, $\Delta\rho$, and $z_{ref}$.
The model and corresponding synthetic data are meant to simulate
with more fidelity the real data application.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/synthetic-crust1-results}
    \caption{
        Inversion results from the CRUST1.0 synthetic data.
        (a) Cross-validation curve used to determine
        the regularization parameter (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}) is found at
        $\mu = 0.0001$ (red triangle).
        (b) Cross-validation results used to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}) in
        $km^2$.
        The minimum (red triangle) is found at $z_{ref} = 30\ km$
        and $\Delta\rho = 350\ kg/m^3$.
        (c) The estimated Moho depth.
        (d) Difference between the CRUST1.0 model depths
        (Fig.~\ref{fig:crust1-data}a)
        and the estimated depths.
        (e) Histogram of the inversion residuals
        (observed minus predicted data).
        (f) Histogram of the differences between
        the synthetic seismic observations (Fig.~\ref{fig:crust1-data}c)
        and the estimated depths.
        (g) The inversion residuals.
        (h) Difference between the seismic and the estimated depths.
    }
    \label{fig:crust1-results}
\end{figure*}

@@

The cross-validation procedure to determine $\Delta\rho$ and $z_{ref}$
requires knowledge of the Moho depth at certain points
($\mathbf{z}_s^o$ in Eq.~\ref{eq:msehyper}),
usually from seismic experiments.
Thus, we must also generate synthetic seismic data about the Moho depth.
We produce such data by interpolating the Moho depth shown in
Fig.~\ref{fig:crust1-data}a on the same 937 geographic coordinates
pinpointed in the data set of \citet{assumpcao2013a}.
The resulting synthetic seismic data is shown in Fig.~\ref{fig:crust1-data}c.

We perform the cross-validation procedures in two parts.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the cross-validation
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depths.
Fig.~\ref{fig:crust1-results} summarizes the results
from both cross-validation runs and the final inversion results.

For the first cross-validation,
we separate the synthetic data (Fig.~\ref{fig:grid_separation}) into
a training set with twice the grid spacing of the original data
($N_{lat} \times N_{lon} = 80 \times 60$)
and a testing set with 14,121 observations.
We run the inversion for 16 different values of $\mu$
equally spaced in a logarithmic scale between $10^{-7}$ and $10^{-2}$.
For each of the 16 estimates we compute the MSE (Eq.~\ref{eq:msemu}),
shown in Fig.~\ref{fig:crust1-results}a as function of $\mu$.
The optimal regularization parameter that minimizes the MSE is $\mu = 10^{-4}$
(red triangle in Fig.~\ref{fig:crust1-results}a).

In the second cross-validation,
we use the estimated value of $\mu$ in all inversions.
We test seven values of $z_{ref}$ from 20 to 35 km with 2.5 km intervals
and seven values of $\Delta\rho$ from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
We run the inversion for every combination of $z_{ref}$ and $\Delta\rho$,
totaling 49 inversions.
Finally, we calculate the Mean Square Error (Eq.~\ref{eq:msehyper})
for each of the 49 estimates
and choose the values of $z_{ref}$ and $\Delta\rho$ that minimize the MSE.
Fig.~\ref{fig:crust1-results}b
shows a colored-contour map of the MSE
with a minimum (marked by the red triangle)
at $z_{ref} = 30\ km$ and $\Delta\rho = 350\ kg/m^3$.

Fig.~\ref{fig:crust1-results}c shows the final solution after both
cross-validation procedures.
The recovered model is smooth, indicating that the cross-validation procedure
was effective in estimating an optimal regularization parameter.
Fig.~\ref{fig:crust1-results}d shows the difference between
the true Moho depths (Fig.~\ref{fig:crust1-data}a)
and the estimated depths (Fig.~\ref{fig:crust1-results}c).
The maximum and minimum differences are, respectively,
9.8 and -8.2 km.
The largest absolute differences are located along the central and northern
Andes, where there is a sharp increase in the true Moho depth
(Fig.~\ref{fig:crust1-data}a).
Positive differences (indicating a too shallow estimate)
appear along the central portion of the Andes,
flanked by regions of negative differences (indicating a too deep estimate)
on the continental and Pacific sides.
Figs.~\ref{fig:crust1-results}e and g show the gravity residuals,
defined as the difference between the observed and predicted gravity data.
The residuals appear normally distributed,
with 0.03 mGal mean and a standard deviation of 4.10 mGal.
The gravity residuals follow a similar, though reversed, pattern
to the differences shown in Fig.~\ref{fig:crust1-results}d.
The largest residuals (in absolute value) are along the Andes,
with the central portion being dominated by negative residuals
and flanked by positive residuals on both sides.
Figs.~\ref{fig:crust1-results}f and h show the differences between
the synthetic seismic data (Fig.~\ref{fig:crust1-data}c)
and the estimated Moho depths.
Once more, the largest differences are concentrated along the Andes,
particularly in the central Andes and near Ecuador and Colombia.
The differences are smaller along the Atlantic coast of South America,
with notable larger differences in a few points of northeastern Brazil
and along the Amazon river.
In general, large residuals are associated with sharp increases in Moho depth.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Gravity and seismic data}

@@

Fig.~\ref{fig:sam-data}b shows the 937 known Moho depths (colored dots) which
were estimated from seismological data by \citet{assumpcao2013a}.
This data set is used in the cross-validation procedure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inversion and cross-validation}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/south-america-cv}
    \caption{
        Cross-validation results for the South American Moho inversion.
        (a) Cross-validation to determine the regularization parameter $\mu$
        (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}),
        shown as a red triangle,
        corresponds to $\mu = 10^{-10}$.
        (b) Cross-validation to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}).
        The minimum (red triangle) is found at $z_{ref} = 35\ km$
        and $\Delta\rho = 400\ kg/m^3$.
    }
    \label{fig:sam-cv}
\end{figure}

As in the CRUST1.0 synthetic data test (section~\ref{sec:crust1}),
we perform the cross-validation in two parts.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the cross-validation
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depth model.

@@

We proceed with the second cross-validation using $\mu = 10^{-10}$ in all
inversions.
We test all combinations of
seven values of $z_{ref}$, from 20 to 40 km with 2.5 km intervals,
and seven values of $\Delta\rho$, from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
Fig.~\ref{fig:sam-cv}b shows a map of the MSE
with respect to the \citet{assumpcao2013a} data set.
The MSE has a minimum, indicated by the red triangle,
at $z_{ref} = 35\ km$ and $\Delta\rho = 400\ kg/m^3$.
The minimum is not as well-defined as for the CRUST1.0 synthetic
(Fig.~\ref{fig:crust1-results}b),
which is expected because $\Delta\rho$ is not homogeneous across all of
South America and the surrounding oceans.

@@

We have developed a computationally efficient gravity inversion method in
spherical coordinates.
Our method extends the Gauss-Newton formulation of Bott's method
\citep{silva2014} to use tesseroids as model elements and Tikhonov
regularization.
The computational efficiency of our method is due to
using Bott's approximation for the Jacobian matrix
(Eq.~\ref{eq:bott-gaussnewton})
and using sparse matrix algorithms for arithmetic operations and the solution
of linear systems.
The Bouguer plate approximation in Eq.~\ref{eq:bott-gaussnewton} is adequate
and the method converges
even when the data are at higher altitudes and the model is not outcropping,
as shown by the applications to synthetic data.
We employ two cross-validation techniques to estimate the hyper-parameters of
the inversion: the regularization parameter, the Moho density-contrast, and the
Normal Earth Moho depth.

@@

The more complex synthetic data test based on CRUST1.0
(Fig.~\ref{fig:crust1-results})
shows that the cross-validation using pointwise Moho depth information
is able to correctly estimate the density-contrast ($\Delta\rho$) and Normal
Earth Moho depth ($z_{ref}$).
This test indicates that the inversion neither correctly estimates Moho depth
nor adequately fits the gravity and pointwise data when sharp variations in
Moho depth occur.
This phenomenon is particularly strong in the region below the Andes.
A likely explanation is that the smoothness regularization
is intrinsically unable to produce sharp variations in Moho depth.
These effects might be mitigated with the use of sharpness-inducing
regularization, like
the weighted smoothness inversion \citep{barbosa1999a},
Cauchy norm regularization \citep{sacchi1996, pilkington2008},
entropic regularization \citep{silva2010},
total variation regularization \citep{martins2011},
or an adaptive mixed smoothness-sharpness regularization \citep{sun2014}.

@@

We are indebted to the developers and maintainers of the open-source
software without which this work would not have been possible.
We thank Dr. Marcelo Assumpção for making available to us the seismological
Moho depth estimates.
L. Uieda was supported by a scholarship from
Coordenação de Aperfeiçoamento de Pessoal de Nível Superior
(CAPES).
V.C.F. Barbosa was supported by a fellowship from
Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq).

>>>>

The Mohorovičić discontinuity (or Moho) that marks the transition from the
crust to the mantle, is studied almost exclusively through indirect geophysical
methods.
The two main geophysical methods used to estimate the depth of the Moho are
seismology, with both natural and controlled sources, and gravimetry.
With the advent of satellite gravimetry missions like GRACE and GOCE,
gravity-derived crustal models can be produced in regional or global scales
\citep[e.g. ][]{reguzzoni2013,vandermeijde2013,vandermeijde2015}.
New spherical harmonic gravity models that use these satellite observation,
like GOCO5S \citep{mayer-guerr2015}, provide almost homogeneous data coverage
in difficult to access regions traditionally poor in terrestrial data.
An example is South America, where seismological and terrestrial gravity data
are traditionally concentrated around urban centers and coastal areas,
resulting in large areas (e.g., forests and mountains) devoid of data.

@@

We present a non-linear gravity inversion to estimate the Moho depth in
a spherical Earth approximation.
Our method is based on the \citet{silva2014} Gauss-Newton formulation of the
method of \citet{bott1960}.
We use tesseroids to discretize the anomalous Moho and the adaptive
discretization algorithm of \citet{uieda2016} for the forward modeling.
The stability of the inversion is achieved through smoothness regularization.
In order to maintain the computational efficiency of Bott's method,
we exploit the sparse nature of all matrices involved in the computations.
We employ a variant of GCV known as hold-out cross-validation \citep{kim2009}
to estimate the regularization parameter.
Additionally, we estimate the density-contrast and reference level
simultaneously in a second validation step using knowledge of the Moho depth at
a few points, similarly to \citet{silva2006} and \citet{martins2010}.
Finally, we apply the proposed method to estimate the Moho depth for South
America using gravity data from the GOCO5S model \citep{mayer-guerr2015} and
the seismological data of \citet{assumpcao2013a}.

@@

We estimate these hyper-parameters in two steps.
First, we assume fixed values for
$z_{ref}$ and $\Delta\rho$ and perform a hold-out cross-validation procedure
\citep{hansen1992} to estimate an optimal value for $\mu$.
Our investigations suggest that the optimal value of $\mu$ does not depend on
the particular values of $z_{ref}$ and $\Delta\rho$ used.
Second, we use the estimated $\mu$ to perform a validation
procedure to estimate $z_{ref}$ and $\Delta\rho$.
The final outcomes of both steps are the values of the three hyper-parameters
and the final estimated Moho depths.

@@

The depth of the Normal Earth Moho ($z_{ref}$)
and the density-contrast of the anomalous Moho ($\Delta\rho$)
are other hyper-parameters of the inversion.
That is, their value influences the final solution
but they are not estimated during the inversion.
Both hyper-parameters cannot be determined from the gravity data alone.
Estimating $z_{ref}$ and $\Delta\rho$ requires
information that is independent of the gravity data,
such as knowledge of the parameters (Moho depths) at certain points.
This information can be used in a manner similar to
the cross-validation described in the previous section.
In this study, we use point estimates of the Moho depth
to determine the optimal values of $z_{ref}$ and $\Delta\rho$.
These points will generally come from seismologic studies,
like receiver functions, surface wave dispersion, and deep refraction
experiments.

@@

We test and illustrate the proposed inversion method
by applying it to two noise-corrupted synthetic data sets.
The first one is generated by a simple Moho model simulating the transition
from a thicker continental crust to a thinner oceanic crust.
This application uses cross-validation to estimate the regularizing parameter
($\mu$) while assuming that the anomalous Moho density-contrast ($\Delta\rho$)
and the Normal Earth Moho depth ($z_{ref}$) are known quantities.
This first test is simplified in order to investigate solely
the efficiency of the inversion and
the cross-validation procedure to estimate $\mu$.
The second data set is generated by a more complex model derived from
the South American portion of the global CRUST1.0 model \citep{laske2013}.
This second application uses cross-validation to estimate $\mu$ and
the validation procedure using synthetic seismological data
to estimate $\Delta\rho$ and $z_{ref}$.
The model and corresponding synthetic data are meant to simulate
with more fidelity the real data application.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/synthetic-crust1-results}
    \caption{
        Inversion results from the CRUST1.0 synthetic data.
        (a) Cross-validation curve used to determine
        the regularization parameter (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}) is found at
        $\mu = 0.0001$ (red triangle).
        (b) Validation results used to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}) in
        $km^2$.
        The minimum (red triangle) is found at $z_{ref} = 30\ km$
        and $\Delta\rho = 350\ kg/m^3$.
        (c) The estimated Moho depth.
        (d) Difference between the CRUST1.0 model depths
        (Fig.~\ref{fig:crust1-data}a)
        and the estimated depths.
        (e) Histogram of the inversion residuals
        (observed minus predicted data).
        (f) Histogram of the differences between
        the synthetic seismic observations (Fig.~\ref{fig:crust1-data}c)
        and the estimated depths.
        (g) The inversion residuals.
        (h) Difference between the seismic and the estimated depths.
    }
    \label{fig:crust1-results}
\end{figure*}

@@

The validation procedure to determine $\Delta\rho$ and $z_{ref}$
requires knowledge of the Moho depth at certain points
($\mathbf{z}_s^o$ in Eq.~\ref{eq:msehyper}),
usually from seismic experiments.
Thus, we must also generate synthetic seismic data about the Moho depth.
We produce such data by interpolating the Moho depth shown in
Fig.~\ref{fig:crust1-data}a on the same 937 geographic coordinates
pinpointed in the data set of \citet{assumpcao2013a}.
The resulting synthetic seismic data is shown in Fig.~\ref{fig:crust1-data}c.

We estimate the three hyper-parameters in two parts.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the validation
procedure with the synthetic seismological data
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depths.
Fig.~\ref{fig:crust1-results} summarizes the results.

For the cross-validation,
we separate the synthetic data (Fig.~\ref{fig:grid_separation}) into
a training set with twice the grid spacing of the original data
($N_{lat} \times N_{lon} = 80 \times 60$)
and a testing set with 14,121 observations.
We run the inversion for 16 different values of $\mu$
equally spaced in a logarithmic scale between $10^{-7}$ and $10^{-2}$.
For each of the 16 estimates we compute the MSE (Eq.~\ref{eq:msemu}),
shown in Fig.~\ref{fig:crust1-results}a as function of $\mu$.
The optimal regularization parameter that minimizes the MSE is $\mu = 10^{-4}$
(red triangle in Fig.~\ref{fig:crust1-results}a).

In the validation using seismological data,
we use the estimated value of $\mu$ in all inversions.
We test seven values of $z_{ref}$ from 20 to 35 km with 2.5 km intervals
and seven values of $\Delta\rho$ from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
We run the inversion for every combination of $z_{ref}$ and $\Delta\rho$,
totaling 49 inversions.
Finally, we calculate the Mean Square Error (Eq.~\ref{eq:msehyper})
for each of the 49 estimates
and choose the values of $z_{ref}$ and $\Delta\rho$ that minimize the MSE.
Fig.~\ref{fig:crust1-results}b
shows a pseudo-color map of the MSE
with a minimum (marked by the red triangle)
at $z_{ref} = 30\ km$ and $\Delta\rho = 350\ kg/m^3$.

Fig.~\ref{fig:crust1-results}c shows the final solution after
cross-validation and validation using seismological data.
The recovered model is smooth, indicating that the cross-validation procedure
was effective in estimating an optimal regularization parameter.
Fig.~\ref{fig:crust1-results}d shows the difference between
the true Moho depths (Fig.~\ref{fig:crust1-data}a)
and the estimated depths (Fig.~\ref{fig:crust1-results}c).
The maximum and minimum differences are, respectively,
9.8 and -8.2 km.
The largest absolute differences are located along the central and northern
Andes, where there is a sharp increase in the true Moho depth
(Fig.~\ref{fig:crust1-data}a).
Positive differences (indicating a too shallow estimate)
appear along the central portion of the Andes,
flanked by regions of negative differences (indicating a too deep estimate)
on the continental and Pacific sides.
Figs.~\ref{fig:crust1-results}e and g show the gravity residuals,
defined as the difference between the observed and predicted gravity data.
The residuals appear normally distributed,
with 0.03 mGal mean and a standard deviation of 4.10 mGal.
The gravity residuals follow a similar, though reversed, pattern
to the differences shown in Fig.~\ref{fig:crust1-results}d.
The largest residuals (in absolute value) are along the Andes,
with the central portion being dominated by negative residuals
and flanked by positive residuals on both sides.
Figs.~\ref{fig:crust1-results}f and h show the differences between
the synthetic seismic data (Fig.~\ref{fig:crust1-data}c)
and the estimated Moho depths.
Once more, the largest differences are concentrated along the Andes,
particularly in the central Andes and near Ecuador and Colombia.
The differences are smaller along the Atlantic coast of South America,
with notable larger differences in a few points of northeastern Brazil
and along the Amazon river.
In general, large residuals are associated with sharp increases in Moho depth.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Gravity and seismological data}

@@

Fig.~\ref{fig:sam-data}b shows the 937 known Moho depths (colored dots) which
were estimated from seismological data by \citet{assumpcao2013a}.
This data set is used in the validation procedure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inversion, cross-validation, and validation using seismological
data}

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{figures/south-america-cv}
    \caption{
        Cross-validation results for the South American Moho inversion.
        (a) Cross-validation to determine the regularization parameter $\mu$
        (Eq.~\ref{eq:goalfunction}).
        The minimum Mean Square Error (Eq.~\ref{eq:msemu}),
        shown as a red triangle,
        corresponds to $\mu = 10^{-10}$.
        (b) Validation to determine
        the reference level ($z_{ref}$) and the density-contrast ($\Delta\rho$).
        The colors represent the Mean Square Error (Eq.~\ref{eq:msehyper}).
        The minimum (red triangle) is found at $z_{ref} = 35\ km$
        and $\Delta\rho = 400\ kg/m^3$.
    }
    \label{fig:sam-cv}
\end{figure}

As in the CRUST1.0 synthetic data test (section~\ref{sec:crust1}),
we estimate the hyper-parameters in two steps.
First, we run the cross-validation to estimate
an optimal regularization parameter ($\mu$).
The starting estimate for all inversions is
60 km depth for all model parameters.
For this cross-validation,
we keep $z_{ref}$ and $\Delta\rho$ fixed to
$20\ km$ and $500\ kg/m^3$, respectively.
Second, we use the estimated $\mu$ to run the validation
using the seismological data of \citet{assumpcao2013a}
to estimate $z_{ref}$ and $\Delta\rho$,
thus obtaining the final estimated Moho depth model.

@@

We proceed with the validation using seismological data
using $\mu = 10^{-10}$ in all inversions.
We test all combinations of
nine values of $z_{ref}$, from 20 to 40 km with 2.5 km intervals,
and seven values of $\Delta\rho$, from 200 to 500 $kg/m^3$
with 50 $kg/m^3$ intervals.
Fig.~\ref{fig:sam-cv}b shows a pseudo-color map of the MSE
with respect to the \citet{assumpcao2013a} data set.
The MSE has a minimum, indicated by the red triangle,
at $z_{ref} = 35\ km$ and $\Delta\rho = 400\ kg/m^3$.
The minimum is not as well-defined as for the CRUST1.0 synthetic
(Fig.~\ref{fig:crust1-results}b),
which is expected because in reality $\Delta\rho$ is not homogeneous across all
of South America and the surrounding oceans.

@@

We have developed a computationally efficient gravity inversion method in
spherical coordinates.
Our method extends the Gauss-Newton formulation of Bott's method
\citep{silva2014} to use tesseroids as model elements and Tikhonov
regularization.
The computational efficiency of our method is due to
using Bott's approximation for the Jacobian matrix
(Eq.~\ref{eq:bott-gaussnewton})
and using sparse matrix algorithms for arithmetic operations and the solution
of linear systems.
The Bouguer plate approximation in Eq.~\ref{eq:bott-gaussnewton} is adequate
and the method converges
even when the data are at higher altitudes and the model is not outcropping,
as shown by the applications to synthetic data.
We employ hold-out cross-validation to estimate the regularization parameter
and a validation procedure using seismological data to estimate the Moho
density-contrast and the Normal Earth Moho depth.

@@

The more complex synthetic data test based on CRUST1.0
(Fig.~\ref{fig:crust1-results})
shows that the validation using pointwise Moho depth information
is able to correctly estimate the density-contrast ($\Delta\rho$) and Normal
Earth Moho depth ($z_{ref}$).
This test indicates that the inversion neither correctly estimates Moho depth
nor adequately fits the gravity and pointwise data when sharp variations in
Moho depth occur.
This phenomenon is particularly strong in the region below the Andes.
A likely explanation is that the smoothness regularization
is intrinsically unable to produce sharp variations in Moho depth.
These effects might be mitigated with the use of sharpness-inducing
regularization, like
the weighted smoothness inversion \citep{barbosa1999a},
Cauchy norm regularization \citep{sacchi1996, pilkington2008},
entropic regularization \citep{silva2010},
total variation regularization \citep{martins2011},
or an adaptive mixed smoothness-sharpness regularization \citep{sun2014}.

@@

We are indebted to the developers and maintainers of the open-source
software without which this work would not have been possible.
We thank Dr. Marcelo Assumpção for providing the seismological Moho depth
estimates.
L. Uieda was supported by a scholarship from
Coordenação de Aperfeiçoamento de Pessoal de Nível Superior
(CAPES).
V.C.F. Barbosa was supported by a fellowship from
Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq).
